Incidence of confirmed COVID-19 cases in care home staff and residents: Cohort study using electronic records in a large UK care home provider (March-June 2020)

### 2020-07-02 manuscript/manuscript.Rmd
manuscript v0.3

---
title: "Incidence of confirmed COVID-19 cases in care home staff and residents: Cohort study using electronic records in a large UK care home provider (March-June 2020)"
# author: "Dutey-Magni, PF, Williams, H, Hayward, A, Shallcross, L"
bibliography: references.bib
csl: elsevier-vancouver.csl
output:
  html_document:
    df_print: paged
    toc: yes
    code_folding: hide
  word_document:
    toc: yes
    reference_docx: template.docx
link-citations: yes
always_allow_html: yes
---

@@

tline <- dplyr::bind_rows(timelines$timeline)
stopifnot(unique(tline$date_end)>=date_max)
tline <- filter(tline, between(date, date_min, date_max)) %>% 
  mutate(home_type = factor(home_type, c("Residential", "Nursing"))) %>% 
  mutate(
    user_type2 = case_when(
      user_type %in% c("General", "Elderly") ~ "General/elderly",
      user_type == "Dementia" ~ "Dementia",
      TRUE ~ NA_character_ 
      ),
    admission_type2 = case_when(
      admission_type %in% c("Continuing Care", "Independent Living") ~ "Continuing care/independent living", 
      TRUE ~ admission_type
    ))

@@

occupancy_interpolated <- resident_days_linear_approx() #%>% 
  # # TODO: currently putting the individual-level occup for 360 because no occ tally is known!
  # # it's got 57 bed including 5 contract beds, so 
  # left_join(transmute(filter(approx_occupancy, home_code == "360"), 
  #                     date, home_code  , occup360 = trunc(occupancy * 57/52)),
  #           by = c("home_code", "date")) %>%
  # mutate(occupancy = if_else(home_code == "360", occup360, occupancy)) %>% 
  # select(-occup360)

home_mean_occupancy <- occupancy_interpolated  %>% 
  filter(between(date, as.Date("2020-01-01"), as.Date("2020-03-14"))) %>%
  group_by(home_code) %>% 
  summarise(mean_occupancy = mean(occupancy, na.rm = T))

@@

desc_homes <- bind_rows(desc_homes, desc_national)

@@

residents_ct <- tibble(date = seq(date_min, date_max_tallies, 1)) %>% 
  left_join(select(new_cases, date, home_code, 
                     new_confirmed_home, new_confirmed_hospital,
                     new_symptomatic_home, new_symptomatic_hospital,
                     new_deaths_home, new_deaths_hospital)) %>% 
  filter(!is.na(home_code)) %>% 
  left_join(distinct(reference_homes, home_code, home_name_datix, postcode, 
                     staff_total, nursing_dementia, nursing_general, 
                     residential_dementia, residential_general, younger_person)) %>% 
  left_join(select(reference_geography, postcode, region_UK)) %>% 
  left_join(occupancy_interpolated) %>% 
  arrange(home_code, date) %>% 
  group_by(home_code) %>% 
  mutate(new_symptomatic = tidyr::replace_na(new_symptomatic_home, 0) + 
           tidyr::replace_na(new_symptomatic_hospital, 0),
         new_confirmed = tidyr::replace_na(new_confirmed_home, 0) + 
           tidyr::replace_na(new_confirmed_hospital, 0),
         new_deaths = tidyr::replace_na(new_deaths_home, 0) # +                                    tidyr::replace_na(new_deaths_hospital, 0)
         ) %>% 
  mutate(I_symptomatic = total_cases_in_home(occupancy, new_symptomatic, date),
         I_confirmed = total_cases_in_home(occupancy, new_confirmed, date)) %>% 
  ungroup()

km_ct_residents_deaths <- residents_ct %>% 
  group_by(date) %>% 
  summarise(
    deaths = sum(new_deaths),
    occupancy = sum(occupancy)
  ) %>% 
  km_ct_estimator(df = ., t = date, d = deaths, 
                  pop = occupancy, subtract_cases = FALSE, overwrite = TRUE) %>% 
  ungroup() %>% 
  mutate(
    St_max = S_t^exp(1.96*S_t_SElog),
    St_min = S_t^exp(-1.96*S_t_SElog),
    event = "Death"
  )

@@

@@

## OUTBREAK HOMES 
desc_residents_outbreak_only <- list()
desc_residents_outbreak_only$outbreak_homes <- home_total_confirmed$home_code[home_total_confirmed$CT_confirmed != 0]

@@

@@

```{r captions}
## TABLES
table_nums <- captioner::captioner(prefix = "Table")
tab_national_descriptives <- table_nums(
  name = "tab_national_descriptives",
  caption = "Total homes, beds, contract beds, and residents within FSHCG and nationwide")
tab_resident_descriptives <- table_nums(
  name = "tab_resident_descriptives",
  caption = paste0("Breakdown of FSHCG private bed residents by care home type, sex, age, region and status on study exit (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
supp_epi_CT <- table_nums(
  name = "supp_epi_CT",
  caption =  paste0("Incidence proportions and rates of SARS-CoV-2 infections in residents and staff according to FSHCG manager counts (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
supp_epi_CT_oubreak_only <- table_nums(
  name = "supp_epi_CT_oubreak_only",
  caption =  paste0("Incidence proportions and rates of SARS-CoV-2 infections among residents of outbreak care homes according to FSHCG manager counts (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
tab_CoxPH_confirmed <- table_nums(
  name = "tab_CoxPH_confirmed",
  caption =  paste0("Factors associated with the hazard of become a confirmed case: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ")"))
tab_CoxPH_death <- table_nums(
  name = "tab_CoxPH_death",
  caption =  paste0("Factors associated with all-cause mortality: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ")"))

@@

fig_KM_CT_residents <- fig_nums(
  name = "fig_KM_CT_residents",
  caption = paste0("Kaplan-Meier estimates of the cumulative incidence of symptomatic cases, confirmed infections and COVID-related deaths in residents according to FSHCG outbreak surveillance reports (", format(date_min_tallies, "%d/%m/%y"), "-", format(date_max_tallies, "%d/%m/%y"), ")")
)

fig_KM_staff <- fig_nums(
  name = 'fig_KM_staff',
  caption = paste0("Kaplan-Meier estimates of the cumulative incidence of symptomatic cases, confirmed infection in care home staff (n=",   formatbm(desc_staff$total_staff),
  ") according to FSHCG outbreak surveillance reports (", format(date_min_tallies, "%d/%m/%y"), "-", format(date_max_tallies, "%d/%m/%y"), ")"))

@@

**Background** COVID-19 has caused substantial morbidity and mortality in care home residents and staff. National surveys suggest infections are declining in the general population in the UK, but outbreaks persist in care homes. Epidemiological data on disease trends in care home staff and residents are scarce. We used administrative data from the Four Seasons Health Care Group (FSHCG), a national care home chain representing 9% of all beds in England, Scotland and Northern Ireland.

**Methods:** Retrospective cohort study using individual-level electronic records from residents (incidents reports) and daily care home counts of infection for both residents and staff. We estimated the incidence proportion and incidence rate of symptoms, confirmed infection, and death in residents and staff, and compare them with findings from a national community survey. We tested a range of risk factors for resident infection and mortality in multivariate Cox proportional hazards models.

**Findings:** An estimated `r formatbm(desc_residents$CT_total_residents)` residents were present in `r dplyr::n_distinct(residents$home_code)` FSHCG homes between `r format(date_min_tallies, "%e %B %Y")` and `r format(date_max_tallies, "%e %B %Y")`. During the period `r formatbm(desc_residents$CT_total_symptomatic)` residents developed COVID-19 symptoms (`r tbrounding(desc_residents$CT_p_symptomatic)`% [95% confidence interval: `r tbrounding(desc_residents$CT_p_symptomatic_lower)`%; `r tbrounding(desc_residents$CT_p_symptomatic_upper)`%]), and `r formatbm(desc_residents$CT_total_confirmed)` (`r tbrounding(desc_residents$CT_p_confirmed)`% [`r tbrounding(desc_residents$CT_p_confirmed_lower)`%; `r tbrounding(desc_residents$CT_p_confirmed_upper)`%]) had an infection confirmed with nasopharyngeal RNA tests. The confirmed infection incidence rate was `r tbrounding(desc_residents$CT_IR_confirmed)` per 100,000 resident-days [`r tbrounding(desc_residents$CT_IR_confirmed_lower)`; `r tbrounding(desc_residents$CT_IR_confirmed_upper)`] across all care homes, including `r desc_residents$CT_zero_case` homes (`r tbrounding( desc_residents$CT_zero_case/desc_residents$total_homes*100)`%) that had not reported a single confirmed case.
Within England, the confirmed infection rate was `r tbrounding(england_results$results$measure[2,"estimate"])` times [`r tbrounding(england_results$results$measure[2,"lower"]) `; `r tbrounding(england_results$results$measure[2,"upper"])`] greater in care homes than in community households. Out of `r formatbm(desc_staff$total_staff)` staff, `r formatbm(desc_staff$CT_confirmed_total)` confirmed infections were counted(
`r tbrounding(desc_staff$CT_confirmed_p)`% [`r tbrounding(desc_staff$CT_confirmed_p_lower)`%; `r tbrounding(desc_staff$CT_confirmed_p_upper)`%]), corresponding to a rate of incidence of 
`r tbrounding(desc_staff$CT_confirmed_IR)` per 100,000 staff-days [`r tbrounding(desc_staff$CT_symptomatic_IR_lower)`; `r tbrounding(desc_staff$CT_symptomatic_IR_upper)`].

Several risk factors such as being male or aged 85 years and over were independently associated with residents' confirmed infection rate. An increase by 10 percentage points in the staff-to-bed ratio was associated with a 25% reduction [19%, 31%] in the rate of confirmed cases. The ratio of occupants to bedrooms measured at the beginning of the epidemic had a strong positive association with the infection rate. Hazards of confirmed infection in homes with a ratio greater than 1.00 were 4·9 times [2·8; 8·4] greater than in homes with a ratio of 0.70 to 0.85.
Mortality was strongly associated with both being male and age.  An increase by 10 percentage points in the staff-to-bed ratio was associated with a 25% reduction [19%, 31%] in the rate of confirmed cases. The ratio of occupants to bedrooms measured at the beginning of the epidemic had a strong positive association with the infection rate. Hazards of confirmed infection in homes with a ratio greater than 1.00 were 4·9 times [2·8; 8·4] greater than in homes with a ratio of 0.70 to 0.85. All-cause mortality was strongly associated with being male, age, but not rates of staffing or occupancy. Adjusted mortality hazard ratios for asymptomatic and symptomatic confirmed cases respectively were 2.4 [1.4; 4.0] and 10 [8.3; 12] compared with residents that were asymptomatic and never tested positive.

@@

Care home residents are extremely vulnerable to infection due to their advanced age and high prevalence of comorbidity, and this is compounded by frequent exposure to infection, for example through contact with infected staff members. At the peak of the pandemic the number of deaths in UK care home residents was threefold higher than the equivalent period in 2019,[@onsDR2020] but estimates of infection incidence and prevalence over this time period are lacking due to delays in establishing universal testing for care home residents and staff.

For example, at the start of the pandemic, testing for SARS-CoV-2 infection testing using reverse transcriptase polymerase chain reaction (RT-PCR) was only routinely available for care home residents or staff who were admitted to hospital. Testing was also performed in care homes that notified an outbreak to Public Health England, but this was limited to a maximum of five test per care home.  Widespread testing for residents first became available on 11 May 2020 through the whole care home testing programme. [@dhscTesting2020] Consequently national estimates of incidence and prevalence based on test results obtained prior to 11 May will substantially underestimate the burden of infection in care home residents.

An important source of epidemiological data on the burden of disease in care homes are outbreak investigations.[@McMichael2020; @Dora2020] A living systematic review[@Salcher-Konrad2020] of outbreak investigations found incidence proportions ranging between 0% and 72% among residents and between 2% and 64% among staff of long term care facilities. Estimates of this nature cannot be generalised due to the unknown proportion of care homes that have not reported any outbreaks.

Information on disease incidence and prevalence, and how this varies by care home is essential to inform the public health response to future waves of infection, and in particular decisions around the need for widespread universal testing in care home staff and residents.

This study obtained data from `r formatbm(dplyr::n_distinct(residents$home_code))` care homes from the Four Seasons Health Care Group (FSHCG), one of the UK's largest independent provider of long-term and respite residential and nursing care. It estimated incidence of SARS-CoV-2 symptoms, diagnoses and all-cause mortality. In addition, it tested and identified individual and organisational risk factors associated with the rate of confirmed infections.

@@

In 2020, FSHCG represented approximately 9% of all beds in England, Scotland and Northern Ireland (`r table_nums("tab_national_descriptives", display = "cite")`). As of 22 June 2020, 90% of FSHCG care homes have participated in the whole care home testing programme.

FSHCG collects administrative data on their private residents through a variety of electronic systems that are primarily used for billing purposes. This includes individual-level data on the dates of care home entry and exit for residents, demographic information, and reports of infection through an incidents management system known as "Datix". Such records do not include residents occupying beds contracted by local authorities as part of block agreements.

@@

```{r FSHCG_map, fig.width=3, fig.height=5, warning=F}
reference_homes_lat_long <- reference_homes %>% 
  dplyr::filter(home_code %in% home_inclusion) %>% 
  dplyr::select(home_code, home_name_datix, postcode) %>% 
  dplyr::left_join(reference_geography, by = "postcode")

care_homes_map <- ggplot(UK_country_region_sf)  +
  geom_sf(size = 0.1) +
  geom_sf_text(aes(label = ctry_reg_abb), fontface = "bold") +
  theme_void() +
  geom_point(data = reference_homes_lat_long, aes(y = latitude, x = longitude),
             color = "#0099B4FF") +
  ggtitle(paste0(length(home_inclusion), " FSHCG care homes"))
# pdf("map.pdf", width = 3, height = 5)
care_homes_map
# dev.off()
```

@@

## Aggregate outbreak data (`r format(date_min_tallies, "%d/%m/%Y")`--`r format(date_max_tallies, "%d/%m/%Y")`)

On 24 March 2020, FSHCG introduced a bespoke reporting system to monitor COVID-19 cases in each care home. This new system requires care home managers to report daily tallies of symptomatic and confirmed infections identified among staff and residents to their FSHCG regional manager.

Daily counts reported by care homes to FSHCG were obtained for every care home:

- new symptomatic cases (residents)
- new confirmed cases in home (residents)
- new confirmed cases in hospital (residents)
- deaths related to COVID-19 (Haydn please confirm what these correspond to: suspected/coroner report/tested positive before death?)
- new symptomatic cases (staff)
- new confirmed cases (staff)
- deaths related to COVID-19 (staff)

@@

These counts include events relating to residents occupying local authority contract beds as well as a minority of residents receiving a different category of support (eg. young people) who were excluded from the study population.

## Individual-level residents data (`r format(date_min, "%d/%m/%Y")`--`r format(date_max, "%d/%m/%Y")`)

Secondary use individual-level data were extracted and pseudonymised by HW. They consisted of two datasets:

- **a residents dataset** containing one record per resident per care home, alongside residents' gender, date of birth, most recent dates of admission and discharge, date of first admission, type of stay (residential/nursing) and care (general, dementia, elderly). This dataset does not include any records for occupants of local authority 'block contract' beds  ($n=$ `r formatbm(sum(desc_homes[["Total contract beds"]], na.rm=T))`).
- **an incidents dataset** containing `r formatbm(desc_before_exclusions$nrow_incidents)` reports filed by care home staff into FSHCG's Datix incidents management system. Mandated report field were: resident forename, surname, care home identifier, incident date/time, and date/time of reporting. In addition, reports can include: date of birth, gender, information on COVID-19 symptoms, tests and test results, resident current location (home/hospital), and death.

HW created a resident index file bridging the two datasets using: the first three letters of the resident's forename; the resident's full surname; and the care home identifier. Clerical review of identifiers generated in this way established that no single identifier was allocated to distinct individuals in the residents dataset. Just `r desc_before_exclusions$missing_id` (`r tbrounding(desc_before_exclusions$missing_id_pct)`%) Datix reports did not contain enough information to generate an identifier and were excluded. Identifiers of a further `r desc_before_exclusions$link_fail` reports (`r desc_before_exclusions$link_fail_pct`%) could not be successfully linked to the residents dataset and were excluded, after clerical review established they could not be matched with a resident. A proportion of these reports may have been filed in error, while others may relate to occupants of local authority beds who are not accounted for in the residents dataset.

@@

@@

## Descriptives

@@

### Outbreak homes

>>>>

---
title: "Incidence of confirmed COVID-19 cases in care home staff and residents: Cohort study using electronic records in a large UK care home provider (March-June 2020)"
# author: "Dutey-Magni, PF, Williams, H, Hayward, A, Shallcross, L"
bibliography: references.bib
csl: elsevier-vancouver.csl
output:
  word_document:
    toc: yes
    reference_docx: template.docx
  html_document:
    df_print: paged
    toc: yes
    code_folding: hide
link-citations: yes
always_allow_html: yes
---

@@

tline <- dplyr::bind_rows(timelines$timeline)
stopifnot(unique(tline$date_end)>=date_max)
tline <- filter(tline, between(date, date_min, date_max)) %>% 
  mutate(home_type = factor(home_type, c("Residential", "Nursing"))) %>% 
  mutate(
    user_type2 = as.factor(case_when(
      user_type %in% c("General", "Elderly") ~ "General/elderly",
      user_type == "Dementia" ~ "Dementia",
      TRUE ~ NA_character_ 
      )),
    admission_type2 = case_when(
      admission_type %in% c("Continuing Care", "Independent Living") ~ "Continuing care/independent living", 
      TRUE ~ admission_type
    )) %>% 
  mutate(
    user_type2 = relevel(user_type2, "General/elderly")
  )

@@

occupancy_interpolated <- resident_days_linear_approx() #%>% 
home_mean_occupancy <- occupancy_interpolated  %>% 
  filter(between(date, as.Date("2020-01-01"), as.Date("2020-03-14"))) %>%
  group_by(home_code) %>% 
  summarise(mean_occupancy = mean(occupancy, na.rm = T))

@@

desc_homes <- bind_rows(desc_homes, desc_national) 
```

@@

residents_ct <- tibble(date = seq(date_min, date_max_tallies, 1)) %>% 
  left_join(select(new_cases, date, home_code, 
                     new_confirmed_home, new_confirmed_hospital,
                     new_symptomatic_home, new_symptomatic_hospital,
                     new_deaths_home, new_deaths_hospital)) %>% 
  filter(!is.na(home_code)) %>% 
  left_join(distinct(reference_homes, home_code, home_name_datix, postcode, 
                     staff_total, nursing_dementia, nursing_general, 
                     residential_dementia, residential_general, younger_person)) %>% 
  left_join(select(reference_geography, postcode, region_UK)) %>% 
  left_join(occupancy_interpolated) %>% 
  arrange(home_code, date) %>% 
  group_by(home_code) %>% 
  mutate(new_symptomatic = tidyr::replace_na(new_symptomatic_home, 0) + 
           tidyr::replace_na(new_symptomatic_hospital, 0),
         new_confirmed = tidyr::replace_na(new_confirmed_home, 0) + 
           tidyr::replace_na(new_confirmed_hospital, 0),
         new_deaths = tidyr::replace_na(new_deaths_home, 0)  +
           tidyr::replace_na(new_deaths_hospital, 0)
         ) %>% 
  mutate(I_symptomatic = total_cases_in_home(occupancy, new_symptomatic, date),
         I_confirmed = total_cases_in_home(occupancy, new_confirmed, date)) %>% 
  ungroup()

km_ct_residents_deaths <- residents_ct %>% 
  group_by(date) %>% 
  summarise(
    deaths = sum(new_deaths),
    occupancy = sum(occupancy)
  ) %>% 
  km_ct_estimator(df = ., t = date, d = deaths, 
                  pop = occupancy, subtract_cases = FALSE, overwrite = TRUE) %>% 
  ungroup() %>% 
  mutate(
    St_max = S_t^exp(1.96*S_t_SElog),
    St_min = S_t^exp(-1.96*S_t_SElog),
    event = "COVID-19-related death"
  )

@@

desc_residents$CT_total_cdeath <-  sum(new_cases$new_deaths_home, na.rm = T) + sum(new_cases$new_deaths_hospital, na.rm = T)
desc_residents$CT_total_cdeath_denom <- sum(km_ct_residents_deaths$n_t)

desc_residents[paste0(
  "CT_p_cdeath", c("", "_lower", "_upper")
)] <- epitools::binom.exact(desc_residents$CT_total_cdeath,
                            desc_residents$CT_total_residents)[
                              c("proportion", "lower", "upper")
                            ] * 100

desc_residents[paste0(
  "CT_IR_cdeath", c("", "_lower", "_upper")
)] <- epitools::pois.exact(desc_residents$CT_total_cdeath,
                            desc_residents$CT_total_cdeath_denom)[
                              c("rate", "lower", "upper")
                            ] * 100000

@@

## OUTBREAK HOMES 
desc_residents_outbreak_only <- list()
desc_residents_outbreak_only$outbreak_homes <- unique(
  home_total_confirmed$home_code[home_total_confirmed$CT_confirmed != 0],
  distinct(filter(new_cases, tidyr::replace_na(new_deaths_hospital, 0) > 0 |
                    tidyr::replace_na(new_deaths_home, 0)>0), home_code)$home_code
)

@@

desc_residents_outbreak_only$CT_total_cdeath <-  sum(new_cases$new_deaths_home, na.rm = T) + sum(new_cases$new_deaths_hospital, na.rm = T)
desc_residents_outbreak_only$CT_total_cdeath_denom <- residents_ct %>% 
  filter(home_code %in% desc_residents_outbreak_only$outbreak_homes) %>% 
  summarise(n = sum(occupancy)) %>% 
  .$n

desc_residents_outbreak_only[paste0(
  "CT_p_cdeath", c("", "_lower", "_upper")
)] <- epitools::binom.exact(desc_residents_outbreak_only$CT_total_cdeath,
                            desc_residents_outbreak_only$CT_total_residents)[
                              c("proportion", "lower", "upper")
                            ] * 100

desc_residents_outbreak_only[paste0(
  "CT_IR_cdeath", c("", "_lower", "_upper")
)] <- epitools::pois.exact(desc_residents_outbreak_only$CT_total_cdeath,
                            desc_residents_outbreak_only$CT_total_cdeath_denom)[
                              c("rate", "lower", "upper")
                            ] * 100000

@@

```{r captions}
## TABLES
table_nums <- captioner::captioner(prefix = "Table")
tab_national_descriptives <- table_nums(
  name = "tab_national_descriptives",
  caption = "Total homes, beds, contract beds, and residents within FSHCG and nationwide")
tab_resident_descriptives <- table_nums(
  name = "tab_resident_descriptives",
  caption = paste0("Breakdown of FSHCG private bed residents by care home type, sex, age, region and status on study exit (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
incidence_CT_residents <- table_nums(
  name = "incidence_CT_residents",
  caption =  paste0("Incidence proportions and rates of SARS-CoV-2 infections in residents according to FSHCG manager counts (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
supp_epi_CT_oubreak_only <- table_nums(
  name = "incidence_CT_staff",
  caption =  paste0("Incidence proportions and rates of SARS-CoV-2 infections among staff according to FSHCG manager counts (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
tab_CoxPH_confirmed <- table_nums(
  name = "tab_CoxPH_confirmed",
  caption =  paste0("Factors associated with the hazard of become a confirmed case: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ")"))
tab_CoxPH_death <- table_nums(
  name = "tab_CoxPH_death",
  caption =  paste0("Factors associated with all-cause mortality: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ")"))

@@

fig_study_diagram <- fig_nums(
  name = "fig_study_diagram",
  caption = paste0("Study overview: location of FSHCG care homes and diagram of data sources")
)

fig_KM_CT_residents <- fig_nums(
  name = "fig_KM_CT_residents",
  caption = paste0("Kaplan-Meier estimates of the cumulative incidence of symptomatic cases, confirmed infections and COVID-related deaths in residents according to FSHCG outbreak surveillance counts (", format(date_min_tallies, "%d/%m/%y"), "-", format(date_max_tallies, "%d/%m/%y"), ")")
)

fig_KM_staff <- fig_nums(
  name = 'fig_KM_staff',
  caption = paste0("Kaplan-Meier estimates of the cumulative incidence of symptomatic cases, confirmed infection in care home staff (n=",   formatbm(desc_staff$total_staff),
  ") according to FSHCG outbreak surveillance counts (", format(date_min_tallies, "%d/%m/%y"), "-", format(date_max_tallies, "%d/%m/%y"), ")"))

@@

**Background** COVID-19 has caused substantial morbidity and mortality in care home residents and staff. National surveys suggest infections are declining in the general population in the UK, but outbreaks persist in care homes. We measured the incidence of infections among care home staff and residents, using administrative data from the Four Seasons Health Care Group (FSHCG), a national care home chain representing 9% of all beds in England, Scotland and Northern Ireland.

**Methods:** Retrospective cohort study using individual-level electronic records from residents (incidents reports) and daily care home counts of infection for both residents and staff. We estimated the incidence proportion and incidence rate of symptoms, confirmed infections, and deaths in residents and staff. The rate of confirmed infections was compared with data from a national community survey. We tested a range of risk factors for resident infection and mortality in multivariate Cox proportional hazards models.

**Findings:** An estimated `r formatbm(desc_residents$CT_total_residents)` residents were present in `r dplyr::n_distinct(residents$home_code)` FSHCG homes between `r format(date_min_tallies, "%e %B %Y")` and `r format(date_max_tallies, "%e %B %Y")`. During the period `r formatbm(desc_residents$CT_total_symptomatic)` residents developed COVID-19 symptoms (`r tbrounding(desc_residents$CT_p_symptomatic)`% [95% confidence interval: `r tbrounding(desc_residents$CT_p_symptomatic_lower)`%; `r tbrounding(desc_residents$CT_p_symptomatic_upper)`%]), and `r formatbm(desc_residents$CT_total_confirmed)` (`r tbrounding(desc_residents$CT_p_confirmed)`% [`r tbrounding(desc_residents$CT_p_confirmed_lower)`%; `r tbrounding(desc_residents$CT_p_confirmed_upper)`%]) had an infection confirmed with nasopharyngeal RNA tests. The confirmed infection incidence rate was `r tbrounding(desc_residents$CT_IR_confirmed)` per 100,000 resident-days [`r tbrounding(desc_residents$CT_IR_confirmed_lower)`; `r tbrounding(desc_residents$CT_IR_confirmed_upper)`] across all care homes, including `r desc_residents$CT_zero_case` homes (`r tbrounding( desc_residents$CT_zero_case/desc_residents$total_homes*100)`%) which had not reported a single confirmed infection or death.
Within England, the confirmed infection rate was `r tbrounding(england_results$results$measure[2,"estimate"])` times [`r tbrounding(england_results$results$measure[2,"lower"]) `; `r tbrounding(england_results$results$measure[2,"upper"])`] greater in care homes than in community households. Out of `r formatbm(desc_staff$total_staff)` staff, `r formatbm(desc_staff$CT_confirmed_total)` confirmed infections were counted (`r tbrounding(desc_staff$CT_confirmed_p)`% [`r tbrounding(desc_staff$CT_confirmed_p_lower)`%; `r tbrounding(desc_staff$CT_confirmed_p_upper)`%]), corresponding to a rate of incidence of 
`r tbrounding(desc_staff$CT_confirmed_IR)` per 100,000 staff-days [`r tbrounding(desc_staff$CT_symptomatic_IR_lower)`; `r tbrounding(desc_staff$CT_symptomatic_IR_upper)`].

Several risk factors were independently associated with residents' confirmed infection hazards.
Males and those aged 85 years had significantly greater hazards of confirmed infection. Homes with a ratio of occupants to bedrooms (measured at the beginning of the epidemic) in excess of 1.00 were found to have confirmed infections hazards 10 times [7·0; 15] greater than homes with a ratio of 0.70 to 0.85. Conversely, an increase by 10 percentage points in the staff-to-bed ratio was associated with a 23% reduction [18%, 29%] in the hazards of confirmed infections.

All-cause deaths both within homes and in hospitals were recorded for occupants of private beds. 
The all-cause mortality proportion was 
`r tbrounding(desc_residents$DAT_p_death)`% [`r tbrounding(desc_residents$DAT_p_death_lower)`%; `r tbrounding(desc_residents$DAT_p_death_upper)`%], with a confirmed infection case-fatality rate of
`r tbrounding(desc_residents$DAT_CFR)`% [`r tbrounding(desc_residents$DAT_CFR_lower)`%; `r tbrounding(desc_residents$DAT_CFR_upper)`%]. All-cause mortality was strongly associated with both older age and being male, but not with rates of staffing or occupancy. Adjusted mortality hazard ratios for asymptomatic and symptomatic confirmed cases respectively were 2·1 [1·3; 3·6] and 8·5 [7·1; 10] compared with asymptomatic residents who never tested positive for SARS-CoV-2.

@@

Care home residents are vulnerable to infections due to their advanced age and high prevalence of comorbidity. This is compounded by frequent exposure to infections, for example through contact with infected staff members. At the peak of the pandemic the number of deaths in UK care home residents was threefold higher than the equivalent period in 2019.[@onsDR2020]

Information on disease incidence and prevalence, and how they vary across care homes is essential to inform the public health response to future waves of infection, and in particular decisions around the need for widespread universal testing in care home staff and residents. Yet, estimates of infection incidence and prevalence over this time period are lacking due to delays in establishing universal testing for care home residents and staff.

For example, at the start of the pandemic, testing for SARS-CoV-2 infection testing using reverse transcriptase polymerase chain reaction (RT-PCR) was only routinely available for care home residents or staff who were admitted to hospital. Testing was also performed in care homes that notified an outbreak to Public Health England, but this was limited to a maximum of five test per care home. Widespread testing for residents only became available on 11 May 2020 through the whole care home testing programme.[@dhscTesting2020] As a result, national estimates of incidence and prevalence based on test results obtained prior to 11 May substantially underestimate the burden of infection in care home residents.

Outbreak investigations constitute an important source of epidemiological data on the burden of disease in care homes.[@McMichael2020; @Dora2020] A living systematic review[@Salcher-Konrad2020] of outbreak investigations found incidence proportions ranging between 0% and 72% among residents and between 2% and 64% among staff of long term care facilities. Despite this, estimates of this nature cannot be generalised due to the unknown proportion of care homes that have not reported any outbreaks.

This study obtained data from `r formatbm(dplyr::n_distinct(residents$home_code))` care homes from the Four Seasons Health Care Group (FSHCG), one of the UK's largest independent provider of long-term and respite residential and nursing care. It estimated incidence of SARS-CoV-2 symptoms, diagnoses and all-cause mortality across all homes. In addition, it tested and identified individual and organisational risk factors associated with the rate of confirmed infections and deaths.

@@

In 2020, FSHCG counted 9,568 beds, representing 9% of all beds in England, Scotland and Northern Ireland (`r table_nums("tab_national_descriptives", display = "cite")`). Under 9% of these beds were contracted to local authorities for their statutory social care responsibilities.

FSHCG collects administrative data on their private residents through a variety of electronic systems that are primarily used for billing purposes. This includes individual-level data on the dates of care home entry and exit for residents, demographic information, and reports of infection through an incidents management system known as "Datix". Such records do not include residents occupying beds contracted by local authorities as part of block agreements, but home-level aggregate information was nevertheless available on all residents (`r fig_nums("fig_study_diagram", display = "cite")`).

@@

```{r FSHCG_map, fig.width=3, fig.height=5, warning=F}
# reference_homes_lat_long <- reference_homes %>% 
#   dplyr::filter(home_code %in% home_inclusion) %>% 
#   dplyr::select(home_code, home_name_datix, postcode) %>% 
#   dplyr::left_join(reference_geography, by = "postcode")
#  
# care_homes_map <- ggplot(UK_country_region_sf)  +
#   geom_sf(size = 0.1) +
#   geom_sf_text(aes(label = ctry_reg_abb), fontface = "bold") +
#   theme_void() +
#   geom_point(data = reference_homes_lat_long, aes(y = latitude, x = longitude),
#              color = "#0099B4FF", size = .5) + # shape = 3
#   ggtitle(paste0(length(home_inclusion), " FSHCG care homes"))
# cairo_ps("map.ps", width = 3, height = 5)
# care_homes_map
# dev.off()
```

```{r, fig.width=10, fig.height=5}
knitr::include_graphics("study_diagram.svg")
```

@@

`r fig_nums("fig_study_diagram")`

## Aggregate outbreak surveillance counts (`r format(date_min_tallies, "%d/%m/%Y")`--`r format(date_max_tallies, "%d/%m/%Y")`)

On 24 March 2020, FSHCG introduced a bespoke reporting system to monitor COVID-19 cases in each care home. This new system requires care homes to report daily tallies of new cases identified among staff and residents to their FSHCG regional manager. Daily counts were obtained for every care home:

- new symptomatic cases (residents)
- newly confirmed infection in home (residents)
- newly confirmed infection in hospital (residents)
- deaths related to COVID-19 (residents previously confirmed or with a death attributed to SARS-CoV-2 by coroner)
- symptomatic cases (staff)
- newly confirmed cases (staff)

@@

These aggregate data related to private residents, residents occupying local authority contract beds, as well as a minority of residents receiving a different category of support (eg. young people) whose individual-level records were not extracted.

## Individual-level residents records (`r format(date_min, "%d/%m/%Y")`--`r format(date_max, "%d/%m/%Y")`)

Secondary use individual-level data for private residents receiving general and elderly care were extracted and pseudonymised by HW. They consisted of two datasets:

- **a residents dataset** containing one record per resident per care home, alongside residents' gender, date of birth, most recent dates of admission and discharge, date of first admission, type of stay (residential/nursing) and care (general, dementia, elderly). This dataset does not include any records for occupants of local authority 'block contract' beds  ($n=$ `r formatbm(sum(desc_homes[["Total contract beds"]], na.rm=T))`).
- **an incidents dataset** containing `r formatbm(desc_before_exclusions$nrow_incidents)` reports filed by care home staff into FSHCG's Datix incidents management system. Mandated report field were: resident forename, surname, care home identifier, incident date/time, and date/time of reporting. In addition, reports could record: date of birth, gender, information on COVID-19 symptoms, tests and test results, resident current location (home/hospital), and death.

HW created a resident index file bridging the two datasets using: the first three letters of the resident's forename; the resident's full surname; and the care home identifier. Clerical review of identifiers generated in this way established that no single identifier was allocated to distinct individuals in the residents dataset. Just `r desc_before_exclusions$missing_id` (`r tbrounding(desc_before_exclusions$missing_id_pct)`%) incident records did not contain enough information to generate an identifier and were excluded. Identifiers of a further `r desc_before_exclusions$link_fail` records (`r desc_before_exclusions$link_fail_pct`%) could not be successfully linked to the residents dataset and were excluded, after clerical review established they could not be matched with a resident. A proportion of these records may have been filed in error, while others may relate to occupants of local authority beds who are not accounted for in the residents dataset.

@@

Confirmed infection data in both sources can only be treated as underestimates of the true incidence, due to the gradual increase in nasopharyngeal swab testing across the study period. Whole care home testing[@dhscTesting2020] began on 11 May 2020, meaning some confirmed infection are likely to have been missed before this date. By 22 June 2020, 90% of FSHCG care homes had participated in the whole home testing programme.

@@

## Resident cases

@@
